---
title: "Quiz2"
author: "Aman arora"
date: "2/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

### Q1
```{r}
model_fit = lm(mpg~displacement, data = Auto)
```
```{r}
round(model_fit$coefficients["displacement"],2)
```
### Q2
```{r}
sum_mpg = summary(model_fit)
round(sum_mpg$r.squared,2)
```
###Q3
```{r}
model_fit = lm(mpg~.-name, data = Auto)
round(model_fit$coefficients["displacement"],2)
```

###Q4
```{r}
X = c(1,4,100,1,400,4)
b = c(50,20,.07,35,.01,-10)

Y = X%*%b
Y
```

## Q5
- For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is less than 3.5.

- 
For a fixed value of IQ and GPA, females earn more on average than males provided that the GPA is higher than 3.5.



## Q6

The new estimates for b_2b and b_4b are 7 and 1, respectively; all other coefficients stay the same.



## Q7-Q10

```{r}
a0 = 50
a1 = 20
a2 = .07
a3 = 35
a4 = .01
a5 = -10

```


```{r}
w0 = a0+a3

w1= a1+a5

w2 = a2

w3 = a4

```
```{r}
c(w0,w1,w2,w3)
```


## Q11-Q14

```{r}
m0 = a0
m1 = a1
m2 = a2
m3 = a4

```

```{r}
c(m0,m1,m2,m3)
```

## Q15-Q20

```{r}
b0 = 50
b1 = 20
b2 = .07
b3 = 35
b4 = .01
b5 = -10

```

```{r}
c0 = b0+b3
c1 = b1+b5
c2 = b2
c3 = b0-c0
c4 = b4
c5 = b1-c1
```
```{r}
c(c0,c1,c2,c3,c4,c5)

```


#21

RSS will decrease
Rsquare will increase

#22

LS coef. and R-square
```{r}
test = Auto[1:100,]
model_fit = lm(mpg~.-name, data = test)
test[101:200,] = test 
model_fit_dbl = lm(mpg~.-name, data = test)

```
```{r}
sum1 = summary(model_fit)
sum(sum1$residuals^2)
```

```{r}
sum2 = summary(model_fit_dbl)
sum(sum2$residuals^2)

```

#23

1st (wrong)


The LS coefficients are not unique if columns of the nn-by-pp design matrix \mathbf{X}X are linear dependent. (Confirmed)

The LS coefficients are not unique if pp is bigger than nn. (confirmed)

The LS coefficients are unique only if the predictors are uncorrelated with each other.

2nd (correct)

The LS coefficients are not unique if p is bigger than n. (confirmed)


The LS coefficients are not unique if columns of the nn-by-pp design matrix \mathbf{X}X are linear dependent (confirmed)


The LS coefficients are not unique if a non-intercept predictor takes a constant value



