---
title: "Week 3 - Homework"
author: "Aman Arora - amana4@illinois.edu"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---

## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**Solution**:
```{r}
library("MASS")
cat_model = lm(Hwt~Bwt, data = cats)
result = summary(cat_model)
```

* The null hypothesis is $H_0:\beta1 = 0$ i.e there is no relationship  between cat's heart weight and body weight. Alternative hypothesis is  $H1:\beta1 \ne 0$ i.e there is a relationship  between cat's heart weight and body weight

* The t statistic is: `r result$coefficients[2,3]` 

* The p value is 2.2e-16

* At $\alpha = 0.05$  we can reject the NULL hypothesis. 

* Rejection of NULL hypothesis indicates there is significant evidence for a relationship between between cat's heart weight and body weight


**(b)** Calculate a 95% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.

```{r}
intervals = confint(cat_model, level = .95)

```
95% confidence interval for $\beta_1$ is: (`r intervals[2,1]`, `r intervals[2,2]`)


**(c)** Calculate a 90% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
intervals = confint(cat_model, level = .9)
```

90% confidence interval for $\beta_0$ is: (`r intervals[1,1]`, `r intervals[1,2]`)


**(d)** Use a 90% confidence interval to estimate the mean heart weight for body weights of 2.1 and 2.8 kilograms. Which of the two intervals is wider? Why?

```{r}
new_bwt = data.frame(Bwt = c(2.1,2.8))
result = predict(cat_model, newdata = new_bwt, interval = c("confidence"), level =0.9)
result
```

The interval is wider for body weight 2.1 since its father from mean of body weight hence has greater error.



**(e)** Use a 90% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.

```{r}

new_bwt = data.frame(Bwt = c(2.8,4.2))
result = predict(cat_model, newdata = new_bwt, interval = c("prediction"), level =0.9)
result

```


**(f)** Create a scatterplot of the data. Add the regression line, 95% confidence bands, and 95% prediction bands.


```{r}

Bwt_grid = seq(min(cats$Bwt), max(cats$Bwt), by = 0.01)

hwt_ci_band = predict(cat_model, newdata = data.frame(Bwt =Bwt_grid ), interval = "confidence",level = .95)

hwt_pi_band = predict(cat_model, newdata = data.frame(Bwt =Bwt_grid ), interval = "prediction",level = .95)

plot(Hwt~Bwt, data = cats, ylim = c(min(hwt_pi_band),max(hwt_pi_band)), main = "Heart Weight vs Body Weight of cats", xlab= "Body Weight(kg)", ylab ="Heart Weight(grams)")

abline(cat_model, col = "red")

lines(Bwt_grid,hwt_ci_band[,"lwr"],col ="blue",lwd = 3, lty = 2)
lines(Bwt_grid,hwt_ci_band[,"upr"],col ="blue",lwd = 3, lty = 2)

lines(Bwt_grid,hwt_pi_band[,"lwr"],col ="blue",lwd = 3, lty = 3)
lines(Bwt_grid,hwt_pi_band[,"upr"],col ="blue",lwd = 3, lty = 3)

```


**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
result  = summary(cat_model)$coefficients

beta_hat_1 = result[2,1]
se = result[2,2]

#Hypothesized Beta1
beta_hyp = 4

t = (beta_hat_1- beta_hyp)/se

p = 2*pt(abs(t), df = nrow(cats) -2,lower.tail = FALSE)

```
* The Value of test statistic is: `r t`
* the p value is:`r p` 
* At $\alpha = 0.05$ we fail to reject $H_0$


```{r}

summary(cat_model)$coefficients
```




***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 18760613
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```

**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values

**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.

***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 18760613
set.seed(birthday)
n = 25
x = seq(0, 2.5, length = n)
```

**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.

**(c)** What proportion of these intervals contains the true value of $\beta_1$?

**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?

**(e)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 99% confidence interval. Store the lower limits in a vector `lower_99` and the upper limits in a vector `upper_99`.

**(f)** What proportion of these intervals contains the true value of $\beta_1$?

**(g)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.01$?

***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval

**(b)** After writing the function, run this code:

```{r, eval = FALSE}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```

**(c)** After writing the function, run this code:

```{r, eval = FALSE}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.90)
```


