---
title: "HW6"
author: "Aman Arora"
date: "4/26/2020"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Answer 1

```{r cars}
df = read.csv("C://Users//amana//Downloads//illinimensbb.csv",header=TRUE)
plot(Ht~Pos, data = df)
```

There does seems to a correlation between height and position based on the box plots. With tallest players at center,  shorter players at guard and most middle height player in the front.

## Answer 2 (a)

$\underline{JAGS- Model}$

model {
for (i in 1:length(FGM)) {
FGM[i] ~ dbin(prob[i], FGA[i])
logit(prob[i]) <- betapos[Pos[i]] + betaht*htscaled[i]
FGMrep[i] ~ dbin(prob[i], FGA[i])
}
for (j in 1:max(Pos)) {
betapos[j] ~ dt(0, 0.01, 1)
}
betaht ~ dt(0, 0.16, 1)
}


```{r}
d1 <- list(FGM = df$FGM,FGA = df$FGA, Pos = unclass(df$Pos),htscaled = as.vector(scale(df$Ht, scale=2*sd(df$Ht))))

inits1 <- 
list(list(betapos=c(10,10,10),betaht=10,.RNG.name="base::Wichmann-Hill",.RNG.seed=7),
list(betapos=c(10,10,-10), betaht=-10,.RNG.name="base::Wichmann-Hill",.RNG.seed=8),
list(betapos=c(10,-10,10), betaht=-10,.RNG.name="base::Wichmann-Hill",.RNG.seed=9),
list(betapos=c(10,-10,-10), betaht=10,.RNG.name="base::Wichmann-Hill",.RNG.seed=10))
```

```{r}
library(rjags)
m1 <- jags.model("C://Users//amana//Downloads//FGM.bug", d1, inits1, n.chains=4, n.adapt=2000)
update(m1, 1000) # burn-in
x1 <- coda.samples(m1, c("betapos","betaht"),n.iter=10000)
gelman.diag(x1, autoburnin=FALSE)

```
As we can see  the gelam-Rubin statistic is converging to 1 indicating no
convergence problemeffectiveSize(x1[,1:4])



```{r}
effectiveSize(x1[,1:4])

```
## Answer 2 (b)

```{r}

summary(x1[,1:4])


```

## Answer 2 (c)
```{r}
x2 = as.matrix(x1)

plot(x2[,'betaht'],x2[,'betapos[1]'])

plot(x2[,'betaht'],x2[,'betapos[2]'])

plot(x2[,'betaht'],x2[,'betapos[3]'])

```
There seems to be negative  correlation  between beta[1] (center)/beta[2] (forward) and beta height. Also there is positive correlation between beta[3] (guard) and beta height. 

## Answer 2 (d)

Posterior density of FGM probability of Ayo Dosunmu (at index 4)

```{r}
x3 <- coda.samples(m1, c("betapos","betaht","prob","FGMrep"),n.iter=10000)

```
```{r}
x4 <- as.matrix(x3)
densplot(x3[,c("prob[4]")])

```

## Answer 2 (e)

Posterior probability of forward having a higher probability of making a goal vs guards. 

```{r}
cat("Posterior probability:", mean(x4[,'betapos[2]']>x4[,'betapos[3]']),"\n")

bayes_factor <- mean(x4[,'betapos[2]']>x4[,'betapos[3]'])/mean(x4[,'betapos[2]']< x4[,'betapos[3]'])
cat("Bayes factor:", bayes_factor)

```
Bayes factor indicates strong evidence in favor FGM for forward vs guards.

## Answer 2 (f)

```{r}
probs <- as.matrix(x3)[, paste("prob[",1:nrow(df),"]", sep="")]
FGMrep <- as.matrix(x3)[, paste("FGMrep[",1:nrow(df),"]", sep="")]
Tchi <- numeric(nrow(df))
Tchirep <- numeric(nrow(df))
for(s in 1:nrow(FGMrep)){
 
 Tchi[s] <- sum((df$FGM - df$FGA*probs[s,])^2 /
 (df$FGA*probs[s,]*(1-probs[s,])))
 
 Tchirep[s] <- sum((FGMrep[s,] - df$FGA*probs[s,])^2 /
 (df$FGA*probs[s,]*(1-probs[s,])))
 }
mean(Tchirep >= Tchi)

```
THe p value is close to 0, the discrepency measure of replicated data has very low probability to be above discrepency measure of data  indicating evidence overdispersion.

##Answer 2 (g)(i)

$\underline{JAGS- Model}$
model {
for (i in 1:length(FGM)) {
FGM[i] ~ dbin(prob[i], FGA[i])
logit(prob[i]) <- betapos[Pos[i]] + betaht*htscaled[i] + epsilon[i]
epsilon[i] ~ dnorm(0,1/sigmaepsilon^2)
FGMrep[i] ~ dbin(prob[i], FGA[i])
}
for (j in 1:max(Pos)) {
betapos[j] ~ dt(0, 0.01, 1)
}
betaht ~ dt(0, 0.16, 1)
sigmaepsilon ~ dunif(0,10)
}


```{r}
library(rjags)

inits2 <- 
list(list(betapos=c(10,10,10),betaht=10,sigmaepsilon=0.01,.RNG.name="base::Wichmann-Hill",.RNG.seed=7),
list(betapos=c(10,10,-10), betaht=-10,sigmaepsilon=9,.RNG.name="base::Wichmann-Hill",.RNG.seed=8),
list(betapos=c(10,-10,10), betaht=-10,sigmaepsilon=.01,.RNG.name="base::Wichmann-Hill",.RNG.seed=9),
list(betapos=c(10,-10,-10), betaht=10,sigmaepsilon=9,.RNG.name="base::Wichmann-Hill",.RNG.seed=10))

m1 <- jags.model("C://Users//amana//Downloads//FGM_dis.bug", d1, inits2, n.chains=4, n.adapt=2000)
update(m1, 1000) # burn-in
x5 <- coda.samples(m1, c("betapos","betaht","sigmaepsilon","prob","FGMrep"),n.iter=10000)


gelman.diag(x5, autoburnin=FALSE)

```
##Answer 2 (g)(ii)

```{r}
densplot(x5[,c("sigmaepsilon")])
```
```{r}
x6 = as.matrix(x5)

cat("Posterior probability:", mean(x6[,'betapos[2]']>x6[,'betapos[3]']),"\n")

bayes_factor <- mean(x6[,'betapos[2]']>x6[,'betapos[3]'])/mean(x6[,'betapos[2]']< x6[,'betapos[3]'])
cat("Bayes factor:", bayes_factor)

```
The conclusions for new model based on p value and bayes value is different. The evidence is now less strong vs previous model  that forward players probability of making a field goal is higher than guard position player.  The bayes factor also indicates a positive evidence not strong like previous model.   

## Answer 3(a)

model {
for (i in 1:length(blocks)) 
{
  blocks[i] ~ dpois(lambda[i])
  log(lambda[i]) <- logmin[i] + beta.pos[pos[i]] + beta.height*htscaled[i]
  blocksrep[i] ~ dpois(lambda[i])
}

beta.height ~ dnorm(0, 0.0001)

for (i in 1:max(pos)) {
beta.pos[i] ~ dnorm(0, 0.0001)
}

}


```{r}
d2 <- list(blocks = df$BLK, logmin = log(df$MIN), htscaled = as.vector(scale(df$Ht)), pos = unclass(df$Pos))

inits2 <- 
list(list(beta.pos=c(100,100,100),beta.height=10,.RNG.name="base::Wichmann-Hill",.RNG.seed=1981),
list(beta.pos=c(100,100,-100), beta.height=-10,.RNG.name="base::Wichmann-Hill",.RNG.seed=1982),
list(beta.pos=c(100,-100,100), beta.height=-10,.RNG.name="base::Wichmann-Hill",.RNG.seed=1983),
list(beta.pos=c(100,-100,-100), beta.height=10,.RNG.name="base::Wichmann-Hill",.RNG.seed=1984))
library(rjags)

m2 <- jags.model("C://Users//amana//Downloads//Shots_blocked.bug", d2, inits2, n.chains=4, n.adapt=1000)

update(m2, 1000) # burn-in
x6 <- coda.samples(m2, c("beta.pos","beta.height","blocksrep","lambda"),n.iter=50000)
gelman.diag(x6[,1:19], autoburnin=FALSE)

```

```{r}
effectiveSize(x6[,1:4])

```

## Answer 3(b)


```{r}
summary(x6[,1:4])


```
## Answer 3(c)

```{r}
x7 <- as.matrix(x6)

post.samp <- x7[,"beta.height"]
post.samp.change = exp(post.samp)
quantile(post.samp.change,c(0.025,0.975))

```

Since the lower end of  95 central inerval is greater than 1 (~1.6) this indicates greater height is associated with higher  blocking rate.

## Answer 3(d)

```{r}
lambdas <- as.matrix(x6)[, paste("lambda[",1:nrow(df),"]", sep="")]
blocksrep <- as.matrix(x6)[, paste("blocksrep[",1:nrow(df),"]", sep="")]
Tchi <- numeric(nrow(blocksrep))
Tchirep <- numeric(nrow(blocksrep))

for(s in 1:nrow(blocksrep)){
 Tchi[s] <- sum((df$BLK - lambdas[s,])^2 / lambdas[s,])
 Tchirep[s] <- sum((blocksrep[s,] - lambdas[s,])^2 / lambdas[s,])
}
 mean(Tchirep >= Tchi)


```

The low p values indicates evidence of a problem.


## Answer 3 (e)(i)

```{r}

for (i in 1:nrow(df))
{
  df$p_value[i] <-  mean(blocksrep[,i] > df$BLK[i])
}  

df[,c("Player","p_value")]
```

## Answer 3 (e)(ii)

One of the player with p values less than .05 is Kane, Samba

## Answer 3 (e)(iii)

Some of the players have probability close to 1 - "De La Rosa, Adonis" and "Jones, Tevian" since they blocked less shots than the play time and height would suggests. Hence the replicated values are greater  than actual values in majority of replicates samples. 


