---
title: "Week 8 - Homework"
author: "Aman Arora (amana4@illinois.edu)"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.alin = "center")
```

## Exercise 1 (Writing Functions)

**(a)** Write a function named `diagnostics` that takes as input the arguments:

- `model`, an object of class `lm()`, that is a model fit via `lm()`
- `pcol`, for controlling point colors in plots, with a default value of `grey`
- `lcol`, for controlling line colors in plots, with a default value of `dodgerblue`
- `alpha`, the significance level of any test that will be performed inside the function, with a default value of `0.05`
- `plotit`, a logical value for controlling display of plots with default value `TRUE`
- `testit`, a logical value for controlling outputting the results of tests with default value `TRUE`

The function should output:

- A list with two elements when `testit` is `TRUE`:
    - `p_val`, the p-value for the Shapiro-Wilk test for assessing normality
    - `decision`, the decision made when performing the Shapiro-Wilk test using the `alpha` value input to the function. "Reject" if the null hypothesis is rejected, otherwise "Fail to Reject."
- Two plots, side-by-side, when `plotit` is `TRUE`:
    - A fitted versus residuals plot that adds a horizontal line at $y = 0$, and labels the $x$-axis "Fitted" and the $y$-axis "Residuals." The points and line should be colored according to the input arguments. Give the plot a title. 
    - A Normal Q-Q plot of the residuals that adds the appropriate line using `qqline()`. The points and line should be colored according to the input arguments. Be sure the plot has a title. 

Consider using this function to help with the remainder of the assignment as well.


```{r}
diagnostics <- function(model, pcol = "grey", lcol = "dodgerblue", alpha = 0.05, plotit = TRUE, testit = TRUE)
{
  result = list(p_val = 0, decision = "NA")
  if(plotit == TRUE)
  {
    par(mfrow = c(1,2))
    plot(fitted(model), resid(model), xlab = "Fitted", ylab = "Residuals", col = pcol )
    abline(h =0, col = lcol)

    qqnorm(resid(model),col = pcol )
    qqline(resid(model),col = lcol)
    
  }
  
  if(testit == TRUE)
  {
    stest = shapiro.test(resid(model))
    result[["p_val"]] =  stest$p.value
    result[["decision"]] = ifelse(stest$p.value >alpha,"Fail to Reject", "Reject")
    return (result)
  }
  
}

```


**(b)** Run the following code.

```{r}
set.seed(40)

data_1 = data.frame(x = runif(n = 30, min = 0, max = 10),
                    y = rep(x = 0, times = 30))
data_1$y = with(data_1, 2 + 1 * x + rexp(n = 30))
fit_1 = lm(y ~ x, data = data_1)

data_2 = data.frame(x = runif(n = 20, min = 0, max = 10),
                    y = rep(x = 0, times = 20))
data_2$y = with(data_2, 5 + 2 * x + rnorm(n = 20))
fit_2 = lm(y ~ x, data = data_2)

data_3 = data.frame(x = runif(n = 40, min = 0, max = 10),
                    y = rep(x = 0, times = 40))
data_3$y = with(data_3, 2 + 1 * x + rnorm(n = 40, sd = x))
fit_3 = lm(y ~ x, data = data_3)
```

```{r}
diagnostics(fit_1, plotit = FALSE)$p_val
diagnostics(fit_2, plotit = FALSE)$decision
diagnostics(fit_1, testit = FALSE, pcol = "black", lcol = "black")
diagnostics(fit_2, testit = FALSE, pcol = "grey", lcol = "green")
diagnostics(fit_3)
```

***

## Exercise 2 (Prostate Cancer Data)

For this exercise, we will use the `prostate` data, which can be found in the `faraway` package. After loading the `faraway` package, use `?prostate` to learn about this dataset.

```{r, message = FALSE, warning = FALSE}
library(faraway)
```

**(a)** Fit an additive multiple regression model with `lpsa` as the response and the remaining variables in the `prostate` dataset as predictors. Report the $R^2$ value for this model.

```{r}
model_fit =  lm(lpsa~., data = prostate)
summary(model_fit)$r.squared

```
  
The $R^2$ value for this model is 0.6548

**(b)** Check the constant variance assumption for this model. Do you feel it has been violated? Justify your answer.

```{r}
library(lmtest)
diagnostics(model_fit)
bptest(model_fit)

```
  
* The constant variance  assumption seems to holding based on visual inspection of the residual vs fitted graph and high p value of 0.3 from  Breusch-Pagan test 

**(c)** Check the normality assumption for this model. Do you feel it has been violated? Justify your answer.

```{r}

diagnostics(model_fit)

```

* The normal assumption seems to holding based on visual inspection of the QQ plot and supported by large p value from Shapiro-Wilk test.

**(d)** Check for any high leverage observations. Report any observations you determine to have high leverage.

```{r}
high_lev_idx = hatvalues(model_fit) > 2*mean(hatvalues(model_fit))

prostate [high_lev_idx,]

```
  
  * There are 5 observations with high leverage as shown above 

**(e)** Check for any influential observations. Report any observations you determine to be influential.

```{r}
influ_idx =  cooks.distance(model_fit) > 4/length(cooks.distance(model_fit))
prostate[influ_idx,]

```

  * There are 7 influential observations  as shown above 

**(f)** Refit the additive multiple regression model without any points you identified as influential. Compare the coefficients of this fitted model to the previously fitted model.

```{r}
#Find indices of rows that do not have influential observations
retain_idx =  cooks.distance(model_fit) < 4/length(cooks.distance(model_fit))

#Fit a new model without influential observations
model_new = lm(lpsa ~ ., data = prostate[retain_idx,])

coef(model_fit)
coef(model_new)


```

The new coefficients after removing influential observations are significantly different than original coefficients. 


**(g)** Create a data frame that stores the observations that were "removed" because they were influential. Use the two models you have fit to make predictions with these observations. Comment on the difference between these two sets of predictions.


```{r}
#new data frame with influential observation 
new_data1 = prostate[influ_idx,]

#predict with original model
result = predict(model_fit, newdata = new_data1)

#predict with model with influential observation removed
result_new = predict(model_new, newdata = new_data1)

#The actual response values
real = prostate[influ_idx, "lpsa"]

#Plot all the predictions and real data
plot(result, col = "blue", ylim = c(0,6), pch =1, ylab = "Prediction")

points(result_new,col = "red", pch =2)

points(real,col = "green", pch =3)
grid()

legend("bottomright", c("original model", "inf. pt. removed ","actual data"), col= c("blue","red","green"), pch = 1:3)


```
  
  * We can see from the plot the prediction with original model are closer to actual value than the model which has the influential points removed.

***

## Exercise 3 (Why Bother?)

**Why** do we care about violations of assumptions? One key reason is that the distributions of the parameter estimators that we have used are all reliant on these assumptions. When the assumptions are violated, the distributional results are not correct, so our tests are garbage. **Garbage In, Garbage Out!**

Consider the following setup that we will use for the remainder of the exercise. We choose a sample size of 50.

```{r}
n = 50
set.seed(420)
x_1 = runif(n, 0, 5)
x_2 = runif(n, -2, 2)
```

Consider the model,

\[
Y = 4 + 1 x_1 + 0 x_2 + \epsilon.
\]

That is,

- $\beta_0$ = 4
- $\beta_1$ = 1
- $\beta_2$ = 0

We now simulate `y_1` in a manner that does **not** violate any assumptions, which we will verify. In this case $\epsilon \sim N(0, 1).$

```{r}
set.seed(83)
library(lmtest)
y_1 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = 1)
fit_1 = lm(y_1 ~ x_1 + x_2)
bptest(fit_1)
```

Then, we simulate `y_2` in a manner that **does** violate assumptions, which we again verify. In this case $\epsilon \sim N(0, \sigma = |x_2|).$

```{r}
set.seed(83)
y_2 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = abs(x_2))
fit_2 = lm(y_2 ~ x_1 + x_2)
bptest(fit_2)
```

**(a)** Use the following code after changing `birthday` to your birthday.

```{r}
num_sims = 2500
p_val_1 = rep(0, num_sims)
p_val_2 = rep(0, num_sims)
birthday = 19081014
set.seed(birthday)
```

Repeat the above process of generating `y_1` and `y_2` as defined above, and fit models with each as the response `2500` times. Each time, store the p-value for testing,

\[
\beta_2 = 0,
\]

using both models, in the appropriate variables defined above. (You do not need to use a data frame as we have in the past. Although, feel free to modify the code to instead use a data frame.)

**(b)** What proportion of the `p_val_1` values is less than 0.01? Less than 0.05? Less than 0.10? What proportion of the `p_val_2` values is less than 0.01? Less than 0.05? Less than 0.10? Arrange your results in a table. Briefly explain these results.

***

## Exercise 4 (Corrosion Data)

For this exercise, we will use the `corrosion` data, which can be found in the `faraway` package. After loading the `faraway` package, use `?corrosion` to learn about this dataset.

```{r, message = FALSE, warning = FALSE}
library(faraway)
```

**(a)** Fit a simple linear regression with `loss` as the response and `Fe` as the predictor. Plot a scatterplot and add the fitted line. Check the assumptions of this model.

```{r}
fitted_model = lm(loss~Fe, data = corrosion )

plot(loss~Fe, data = corrosion)
abline(fitted_model, col = "red")

#Evaluate the assumptions
diagnostics(fitted_model)
bptest(fitted_model)

```
  
* The assumption of constant variance seems to hold from  fitted vs residual plot and high p value from Breusch-Pagan test

* The assumption of normal residuals seems to hold from  Q-Q plot and high p value from Shapiro-Wilks test

**(b)** Fit higher order polynomial models of degree 2, 3, and 4. For each, plot a fitted versus residuals plot and comment on the constant variance assumption. Based on those plots, which of these three models do you think are acceptable? Use a statistical test(s) to compare the models you just chose. Based on the test, which is preferred? Check the normality assumption of this model. Identify any influential observations of this model.


```{r}
library(knitr )
model2 = lm(loss~ Fe + I(Fe^2), data = corrosion )
model3 = lm(loss~ Fe + I(Fe^2) + I(Fe^3), data = corrosion )
model4 = lm(loss~ Fe + I(Fe^2)+ I(Fe^3)+ I(Fe^4), data = corrosion )

name_model = c("2nd degree", "3rd degree", "4th Degree" )
p_val  = c(diagnostics(model2)$p_val, diagnostics(model3)$p_val, diagnostics(model4)$p_val)

result = data.frame("Model Type" = name_model, "Shaprio(p value) " =  p_val ) 

```

`r kable(result,"markdown",caption ="Model Comparison")`  


* Based on the plots and shaprio wilk test  the model with degree 3 seems to have the residuals with  variance that seems constant and highest p value. 

```{r}
bptest(model3)
```


* Based on the normal QQ plots and high p value in  BP test  the model with degree 3 seems to hold the normality assumption as well. 


```{r}
inf_idx =  cooks.distance(model3) >  4/length(cooks.distance(model3))
corrosion[inf_idx,]

```
  
  
* No influential observations were found in the data for the model with degree 3.

***

## Exercise 5 (Diamonds)

The data set `diamonds` from the `ggplot2` package contains prices and characteristics of 54,000 diamonds. For this exercise, use `price` as the response variable $y$, and `carat` as the predictor $x$. Use `?diamonds` to learn more.

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
```

**(a)** Fit a linear model with `price` as the response variable $y$, and `carat` as the predictor $x$. Return the summary information of this model.

```{r}
model1 = lm(price ~ carat, data = diamonds )
summary(model1)

```

**(b)** Plot a scatterplot of price versus carat and add the line for the fitted model in part **(a)**. Using a fitted versus residuals plot and/or a Q-Q plot, comment on the diagnostics. 


```{r}
plot(price ~ carat, data = diamonds)
abline(model1, col = "red")
diagnostics(model1, testit = FALSE)
```
  
  * From Residuals vs fitted plot we do not see constant variance assumption is appropriate for this model and data.
  * From Q-Q plot we see constant normality  assumption does not  look valid. 

**(c)** Seeing as the price stretches over several orders of magnitude, it seems reasonable to try a log transformation of the response. Fit a model with a logged response, plot a scatterplot of log-price versus carat and add the line for the fitted model, then use a fitted versus residuals plot and/or a Q-Q plot to comment on the diagnostics of the model.

```{r}
qplot(price, data = diamonds, bins = 30)
```
```{r}
model2 = lm(log(price) ~ carat, data = diamonds)

plot(diamonds$carat, log(diamonds$price))
abline(model2, col = "red")

diagnostics(model2, testit = FALSE)

```
  
   
  * From Residuals vs fitted plot we do not see constant variance assumption is appropriate for this model and data.
  * From Q-Q plot we see constant normality  assumption does not  look valid. 

**(d)** Try adding log transformation of the predictor. Fit a model with a logged response and logged predictor, plot a scatterplot of log-price versus log-carat and add the line for the fitted model, then use a fitted versus residuals plot and/or a Q-Q plot to comment on the diagnostics of the model.

```{r}
model3 = lm(log(price) ~ log(carat), data = diamonds)

plot(log(diamonds$carat), log(diamonds$price))
abline(model3, col = "red")

diagnostics(model3, testit = FALSE)


```
   
   
 * From Residuals vs fitted plot we do see constant variance assumption is seems appropriate for this model and data vs previous models although not pefect.
 
 * From Q-Q plot we see constant normality  assumption does look valid vs previous models although not perfect. 


**(e)** Use the model from part **(d)** to predict the price (in dollars) of a 3-carat diamond. Construct a 99% prediction interval for the price (in dollars).

```{r}
newdata1 = data.frame(carat = 3)
result = predict(model3, newdata = newdata1, interval = "prediction", level = 0.99)
result_dollars = exp(result)

```
  
  99% prediction interval of price for a 3-carat diamond is (14959, 57894)

